---
title: "Analysis of my reading journey"
author: "Phan Nguyen"
date: "12/24/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)  # visualization
library(lubridate) #manipulate date
library(knitr) # table

```
# Introduction

Hello everyone, welcome to an Analysis of my reading journey. A little bit of background: I started reading again in 2018, after a long time of not touching books. The last book I had read before was when I was in 8th grade :D. 

Since I started reading in 2018, I've been keeping track of the books I read, rating and reviews on Goodreads, and after two years, I had this small set of data available.I came across it the other day, when Goodreads showed me some stats on their website. 

Moreover, I recently started learning data science on my own, therefore I am so thrilled to analyze this data. My purpose of this analysis is to answer some few questions:
1. How much books I've read over the years?
2. Are there correlation between time of year and the amount of reading ?
3. Do I give better rating than other people do?
4. What is my favourite genre? 

I am new to this, so if you have any comments or suggestions, please let me know!

LET'S DIVE IN!

## The Goodreads Data
First, let grab our data
```{r}
data <- read.csv("C:\\Users\\nguyen_phan\\Downloads\\goodreads_library_export.csv",encoding="UTF-8",header=TRUE,stringsAsFactors = FALSE)

```

```{r}
dim(data)
```
There are 57 books in this data. This include books I read, currently am reading, and going to read. This is a small data set, however I believe there are some interesting insights in this one.
Each of this book has 31 variables to describe it. 

This is a sample of the data set. We have name of book, author, my rating, other people' rating, number of pages, date read, date added.
```{r echo=FALSE}
#create a table
kable(data[1:5,][c(2,3,8,9,11,12,15,16,19)])
```

First, I will convert some variables to its proper data type.
```{r echo=FALSE}
#turn variable into date
data$Date.Read <- as.Date(data$Date.Read)
data$Date.Added <- as.Date(data$Date.Added)
```

## Books
```{r}
shelf1 <- as.data.frame(table(data$Exclusive.Shelf))
colnames(shelf1)[1] <- "Shelf"
kable(shelf1)
```

I have read 51 books so far.

We are only interested in looking at books I have read, therefore we will filter read books.

And also make make some variables less mouthful.

```{r}
# filter read books
rdata<-data %>% filter(Exclusive.Shelf=='read')
#easier to see
rdata <- as_tibble(rdata)

#make this variable sh
names(rdata)[names(data)=='Original.Publication.Year'] <- "opy"
```

I am interested in seeing read books by year
```{r}
# create shelf 2, dataframe split books read by year
shelf2 <- rdata %>% group_by(year(Date.Read)) %>% select('year(Date.Read)') 
shelf2 <- as.data.frame(table(shelf2[,1]))
shelf2 <- mutate(shelf2,pct_change=round((Freq/lag(Freq)-1)*100,2))
colnames(shelf2) <- c("Year","Books Read","Percent Change")
```
```{r}
#create a table
kable(shelf2)
```

```{r echo=FALSE}
# ggplot(shelf2, aes(x=Year,y=Freq))+geom_col(fill="aquamarine4")+
#   geom_text(aes(label=Freq), vjust=-0.5)+
#   labs(y="Number of Books",title="Books Read per Year")+
#   ylim(0,25)
```
There was a big jump between 2018 and 2019 in terms of book read (111%). However, there was no significant change from 2019 to 2020 (21%). I'm reading more books each year, but at a decreasing rate.

Here are the oldest and newest books
```{r results="hold"}
#create a table with row as the title of the newest books, remove missing values
table <- tibble(rdata$Title[rdata$opy==max(na.omit(rdata$opy))&is.na(rdata$opy)==FALSE])
table[,2] <- rep(max(na.omit(rdata$opy),nrow(table)))
colnames(table) <- c("Newest books","Year")
kable(table)

#create a table with row as the title of the oldest books, remove missing values
table <- tibble(rdata$Title[rdata$opy==min(na.omit(rdata$opy))&is.na(rdata$opy)==FALSE])
table [,2] <- rep(min(na.omit(rdata$opy)),nrow(table))
colnames(table) <- c("Oldest books","Year")
kable(table)
```

Here are the thickest book and the least thickest book
```{r warning=FALSE, message=FALSE}
table1 <- tibble(rdata$Title[which.max(rdata$Number.of.Pages)])
max(rdata$Number.of.Pages,na.rm = TRUE)
rdata$Title[which.min(rdata$Number.of.Pages)]
min(rdata$Number.of.Pages,na.rm=TRUE)
```


## Speed
Another information I want to know is how fast do I finish a book? 
I will create a variable in rdata, to calculate the speed.
Speed will equal to Number of pages divided by Days Difference or **datediff**.

**datediff** is calculated by taking Date Read minus Date Added
```{r}
# days to finish a book
rdata$datediff <- difftime(rdata$Date.Read,rdata$Date.Added,units = "day")
rdata$datediff
```
The disadvantage of Goodreads export data is that it only allows **Date Added** and **Date Read**. **Date Added** is when I added the book to my shelf, and Date Read is when I claimed to finish it. Normally, **Date Read** will be after **Date Added**, if I added on the day I started reading it, which results in a positive **datediff**. 

There are instances where I updated a book after I finished it, therefore **Date Read** will be before **Date Added**. The negative represents those cases. 

The 0s are the cases where I added the book right on the day I finished it.

There is another column called Date Started on Goodreads, but they don't allow you to export it. 

```{r}
#count datediff which smaller than 1
aggregate(rdata$datediff,list(rdata$datediff<1), FUN=length)
```

I will remove the 0 and the negative to calculate speed. 17 books will be removed

```{r}
rdata$speed  <- rdata$Number.of.Pages/as.integer(rdata$datediff)
speedat <- rdata%>% filter(speed>0&is.na(rdata$speed)==FALSE&speed!=Inf)
speedat$speed
mean(speedat$speed)
```
My average reading speed is 33.31 pages/day

Let's look at it through year
```{r}
speedat %>% group_by(year(Date.Read))%>%summarise(avgspeed = mean(speed,na.rm=TRUE))
```

Average speed in 2018 somehow is the highest, despite the low number of books. Is it because of the number of pages per book?

```{r}
kable(speedat %>% group_by(year(Date.Read))%>%summarise(avgspeed = mean(speed,na.rm=TRUE),avgpages = mean(Number.of.Pages,na.rm=TRUE),n=n()) %>% mutate(truen=shelf2$`Books Read`) %>% mutate(removed=truen-n))
```

The average number of pages in 2018 is actually higher than 2019 and 2020. 
From what I understand, I read less books in 2018, but in a short window. In 2020, I read more books, more frequently, but at slower pace. 

The removed data also play a part here. There are 3 books removed for 2018, 9 for 2019 and 5 for 2020. Even though 2020 has double the books of 2018, it might be that the books which are removed in 2020 have a really high number of pages.

Or I just read super fast in a short window in 2018. 

## Number of Pages
We will look at number of pages I've read. Books have different number of pages.
```{r}
datagraph <- data %>% mutate(month=floor_date(Date.Read,unit="month")) %>% group_by(month) %>% mutate(sumpages=sum(Number.of.Pages,na.rm = TRUE))
```

```{r echo=FALSE, warning=FALSE}
datagraph %>%
  ggplot(aes(x=month,y=sumpages))+
  geom_bar(position="dodge",
           stat="identity",
           fill="#00AFBB",width = 20) +
  geom_text(aes(label=sumpages),vjust=-.5,size=2.5)+
  labs(x="Months",y="Pages Read",title="Pages Read through time")+
  theme(plot.title = element_text(hjust = 0.5))+
  ylim(0,1500)+
  scale_x_date(breaks="month",
               labels = function(x) ifelse(year(x)%in%c(2017,max(year(datagraph$month))),"", format(x,"%b %Y")), 
               limits=as.Date(c('2017-12-15', '2020-12-15')))+
  theme(axis.text.x = element_text(angle=60, hjust = 1))
```

As you can see, I'm not a consistent reader :D.


```{r echo=FALSE, warning=FALSE}
# create a year vector from the data
years <- as.integer(unique(year(data$Date.Read)))
years <- na.omit(years)
years <- sort(years, decreasing = FALSE)
# create a color vector for graph colors
colors <- c("deepskyblue","dodgerblue4","steelblue")
# create a vector for total number of pages read in a year
totalpages <- rep(0,length(years))
for (i in c(1:length(years))){
  totalpages[i] <- sum(data$Number.of.Pages[year(data$Date.Read)==years[i]],na.rm=TRUE)
}


#loop through years to create chart of pages read in each year
for (i in c(1:length(years))){
    print(datagraph %>% filter(year(Date.Read)==as.integer(years[i])) %>% 
    ggplot(aes(x=month,y=sumpages))+
    geom_bar(position="dodge",
             stat="identity",
             fill=colors[i]) +
    geom_text(aes(label=sumpages),vjust=-.5)+
    labs(x="Months",y="Pages Read",title=paste(totalpages[i]," pages read in ",years[i],sep=""))  +
    theme(plot.title = element_text(hjust = 0.5))+
    ylim(0,1500)+
    scale_x_date(breaks="month",
                 labels = function(x) ifelse(year(x)!=years[i],"" , months(x, TRUE)),
                 limits=as.Date(c(paste(toString(years[i]-1),"-12-15",sep=""),paste(toString(years[i]),"-12-15",sep="")))
    )
    )
    }
                 
```

## Rating
I want to compare my ratings with average book rating
```{r}
#My rating vs average rating
mean(data$My.Rating)
mean(data$Average.Rating)
```

```{r}
#creating a dataset of my rating versus average rating
ratedata<-aggregate(rdata$Average.Rating,list(myrating=rdata$My.Rating),FUN="mean")
ratedata$class <- ratedata$myrating
for (i in c(0:nrow(ratedata))){ratedata[i,3]<- ratedata$myrating[i]}
colnames(ratedata)[2] <-"avg"
ratedata <- gather(ratedata,var,value,myrating,avg)
```


```{r}
#plot it
ggplot(ratedata, aes(x=class,y=value,fill=var))+
  geom_bar(stat='identity',position='dodge')+
  geom_text(aes(label=round(value,2)),vjust=-.5,position=position_dodge(width=0.9))+
  labs(y="Rating",title="Books Rating: Phan's versus Average")+
  ylim(0,6)+
  theme(plot.title = element_text(hjust = 0.5),legend.title=element_blank(),axis.title.x = element_blank(),axis.text.x = element_blank(),axis.ticks.x = element_blank()) +
  scale_fill_discrete(labels = c("Average Rating", "My Rating"))
```


